---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(readxl)
```

Uploading the codes (That is the official police code list, approved by the chief.)

```{r}
codes_classified <- read_excel("Data/Exports/FOR REVIEW/Copy of march_14_codes_official.xlsx", sheet = 2)

codes_classified
```

```{r}
clean_codes_classified <- codes_classified %>%
  select(value_clean, priority, property_crime) %>%
  distinct() %>%
  filter(priority != "NA")

clean_codes_classified
```
```{r}
values_doubled_check <- clean_codes_classified %>%
  group_by(value_clean) %>%
  tally() %>%
  arrange(desc(n))

values_doubled_check
```

Perfect. Everything should be great

```{r}
full_offenses_pd_approved <- clean_offenses %>%
  inner_join(clean_codes_classified, by = "value_clean") %>%
  mutate(year = floor_date(date, unit = "year")) %>%
  mutate(clean_year = paste("y", substr(year, 1, 4), sep = "_")) %>%
  select(-rowid)

full_offenses_pd_approved
```
So, I'm going to use it to look at:

percent of each category of priority of crimes throughout the years
percent of property crimes throughout the years


Quick count of total offenses marked each year: 

```{r}
total_offenses_by_year <- full_offenses_pd_approved %>%
  group_by(clean_year) %>%
  tally()

total_offenses_by_year
```


First, this is just a general analysis of the crimes, what went up, what went down

```{r}
year_count <- full_offenses_pd_approved %>%
  group_by(value_clean, clean_year) %>%
  tally() %>%
  arrange(clean_year) %>%
  pivot_wider(names_from = "clean_year", values_from = "n") %>%
  mutate_all(~replace(., is.na(.), 0)) %>%
  mutate(pchange_21_20 = ((y_2021 - y_2020)/y_2020)*100) %>%
  #mutate(pchange_20_19 = ((y_2020 - y_2019)/y_2019)*100) %>%
  #mutate(pchange_19_18 = ((y_2019 - y_2018)/y_2018)*100) %>%
  #mutate(pchange_18_17 = ((y_2018 - y_2017)/y_2017)*100) %>%
  #mutate(pchange_17_16 = ((y_2017 - y_2016)/y_2016)*100) %>%
  arrange(desc(pchange_21_20)) %>%
  mutate(dif_21_20 = abs(y_2021-y_2020)) %>%
  #filter(dif_21_20 > 15) %>%
  #filter(y_2020 >= 20) %>%
  mutate(perc_of_whole_21 = (y_2021/3159)*100) %>%
  mutate(perc_of_whole_20 = (y_2020/2294)*100)  %>%
  mutate(perc_of_whole_19 = (y_2019/3124)*100)  %>%
  mutate(perc_of_whole_18 = (y_2018/2697)*100)  %>%
  mutate(perc_of_whole_17 = (y_2017/3293)*100)  %>%
  mutate(perc_of_whole_16 = (y_2016/3844)*100)  %>%
  mutate(perc_of_whole_15 = (y_2015/3086)*100)  %>%
  mutate(perc_of_whole_14 = (y_2014/2872)*100)  %>%
  mutate(perc_of_whole_13 = (y_2013/2736)*100)  %>%
  mutate(perc_of_whole_12 = (y_2012/2848)*100)  %>%
  mutate(perc_of_whole_11 = (y_2011/3074)*100)  %>%
  arrange(desc(y_2021)) %>%
  select(value_clean, pchange_21_20, y_2021, y_2020, y_2019, y_2018, y_2017, y_2016, y_2015, y_2014, y_2013, y_2012, y_2011, everything())

year_count
```


```{r}
offenses_of_interest <-  year_count %>%
  filter(grepl("SUSPICIOUS", value_clean) | grepl("ACCIDENT", value_clean)| grepl("OPEN", value_clean)| grepl("BURGLARY", value_clean)| grepl("VEHICLE", value_clean)| grepl("CONSTRUCTION", value_clean)| grepl("DWI", value_clean)| grepl("INTOX", value_clean)| grepl("POSS ", value_clean)| grepl("WEAPON", value_clean)| grepl("NOISE", value_clean)| grepl("ANIMAL", value_clean)| grepl("GRAFFITI", value_clean)| grepl("ASSAULT", value_clean)| grepl("ROBBERY", value_clean)| grepl("STALKING", value_clean)| grepl("MURDER", value_clean)| grepl("WELFARE", value_clean))
  #filter(grepl("LOUD", value_clean))

offenses_of_interest
```
```{r}
#write_csv(offenses_of_interest, "Data/Exports/april18_offenses_of_interest.csv")
```

```{r}
drug_poss <- c("POSS CS", "POSS MARIJU", "POSS / OBTAIN CS SCH III/IV F3", "POSS USE ABUSABLE")
```


```{r}
group_offenses_of_interest <- offenses_of_interest %>%
  mutate(group_name = case_when(
      value_clean == "ASSAULT" |  value_clean == "ASSAULT AGAINST PUBLIC SERVANT" | value_clean == "ASSAULT CAUSES BODILY INJURY" | value_clean == "ASSAULT CAUSES BODILY INJURY:FAMILY MEMBER" | value_clean == "ASSAULT CAUSES BODILY INJURY:FAMILY MEMBER (PRIOR CONV)" | value_clean == "ASSAULT THREAT OR OFFENSIVE TOUCH" | value_clean == "ASSAULT THREAT OR OFFENSIVE TOUCH:FAMILY MEMBER" | value_clean == "ASSAULT/ FAMILY/CHOKING/STRANGULAT ION" ~ "Assault",
       value_clean == "ACCIDENT INVOLVING DAMAGE TO VEHICLE" | value_clean == "ACCIDENT INVOLVING PERSONAL INJURY OR DEATH" | value_clean == "ACCIDENT MOTOR VEHICLE" ~ "Accident",
       value_clean == "AGGRAVATED ASSAULT DEADLY WEAPON" | value_clean == "AGGRAVATED ASSAULT DEADLY WEAPON FAMILY MEMBER" | value_clean == "ATT AGGRAVATED ASSAULT PUBLIC SERVANT" | value_clean == "AGGRAVATED ASSAULT DEADLY WEAPON FAMILY MEMBER SERIOUS BODILY INJURY" | value_clean == "AGGRAVATED ASSAULT PUBLIC SERVANT" | value_clean == "AGGRAVATED ASSAULT SBI" ~ "Aggravated Assault",
      value_clean == "AGGRAVATED SEXUAL ASSAULT" | value_clean == "INDECENT ASSAULT" | value_clean == "SEXUAL ASSAULT" | value_clean == "SEXUAL ASSAULT - STATUTORY RAPE" | value_clean == "SEXUAL ASSAULT ON A CHILD" ~ "Sexual Assault",
      value_clean == "AGGRAVATED ROBBERY" | value_clean == "ALARM ROBBERY" | value_clean == "ROBBERY" ~ "Robbery",
      value_clean == "ALARM BURGLARY" | value_clean == "BURGLARY" | value_clean == "BURGLARY COIN OP MACHINE" ~ "Burglary",
      value_clean == "BURGLARY OF A VEHICLE" | value_clean == "THEFT (OF MOTOR VEHICLE PARTS)" | value_clean == "MOTOR VEHICLE THEFT" | value_clean == "ATT BURGLARY VEHICLE" | value_clean == "THEFT MOTOR VEHICLES" | value_clean == "THEFT OF A MOTOR VEHICLE" | value_clean == "BURGLARY OF A VEHICLE - 2 OR MORE PREV CONVIC" | value_clean == "BURGLARY VEHICLE" ~ "Burglary/Theft of Motor Vehicle/Vehicle Parts",
      value_clean == "CAPITAL MURDER" | value_clean == "MURDER" ~ "Murder",
      value_clean == "ANIMAL BITE / SCRATCH CASE" | value_clean == "ANIMAL COMPLAINT" | value_clean == "ANIMAL IMPOUND (NON-COV)" | value_clean == "ANIMAL NUISANCE" | value_clean == "ANIMAL PROGRAM (TNR)" | value_clean == "COV - ANIMAL AT LARGE FELINE (UNALTERED)" | value_clean == "COV - ANIMAL NUISANCE - CHASE PEDESTRIANS BICYCLES OR VEHICLES" | value_clean == "COV - ANIMAL NUISANCE - DAMAGING PROPERTY OF OTHERS" | value_clean == "COV - ANIMAL NUISANCE - DEFECATE ON PROPERTY OF OTHERS" | value_clean == "COV - ANIMAL NUISANCE - DOG FREE OF RESTRAINT" | value_clean == "COV - ANIMAL NUISANCE - LOUD & OBNOXIOUS NOISE" | value_clean == "COV - ANIMAL NUISANCE - MOLEST PASSERBY" | value_clean == "COV - ANIMAL NUISANCE - TRESPASS ONTO PRIVATE PROPERTY" | value_clean == "COV - ANIMAL RESTRICTIONS" | value_clean == "COV - FAIL TO VACCINATE ANIMAL OLDER THAN 4 MONTHS" | value_clean == "COV - OTHER DOMESTIC ANIMALS: RESTRICTIONS" | value_clean == "CRUELTY TO ANIMALS" ~ "Animal-related",
        value_clean == "COV - NOISE NUISANCE" ~ "Noise",
        value_clean == "DRIVING WHILE INTOXICATED >= 015 MA" | value_clean == "DRIVING WHILE INTOXICATED 2ND" | value_clean == "DRIVING WHILE INTOXICATED 3RD OR MORE" | value_clean == "DRIVING WHILE INTOXICATED MB" | value_clean == "DRIVING WHILE INTOXICATION WITH CHILD PASSENGER" | value_clean == "DWI (JUVENILE)" | value_clean == "DWI / DUI (INCLUDES MINORS)" ~ "DWI",
       
        value_clean == "IMPOUNDED VEHICLE" ~ "Impounded Vehicle",
        value_clean == "OPEN DOOR / WINDOW" ~ "Open Door/Window",
        value_clean == "PUBLIC INTOXICATION" ~ "Public Intoxication",
        value_clean == "STALKING" ~ "Stalking",
        value_clean == "SUSPICIOUS PERSON / VEHICLE / CIRCUMSTANCES" ~ "Suspicious",
        value_clean == "UNLAWFULLY CARRYING A WEAPON" | value_clean == "UNLAWFULLY CARRYING A WEAPON (WEAPON FREE ZONE)" | value_clean == "POSSESSION PROHIBITED WEAPON" ~ "Weapon-Issue",
        value_clean == "WELFARE CHECK" ~ "Welfare Check",
      value_clean == "COV - AUTHORIZED HOURS OF CONSTRUCTION" ~ "Construction",
      grepl("GRAFFITI", value_clean) ~ "Graffiti",
      grepl(paste(drug_poss, collapse="|"), value_clean) ~ "Drug Possession"
      ))%>%
  filter(!is.na(group_name)) %>%
  arrange(group_name)

group_offenses_of_interest
```

```{r}
group_offenses_of_interest_sum <- group_offenses_of_interest %>%
  group_by(group_name) %>%
  summarise("2011" = sum(y_2011), "2012" = sum(y_2012), "2013" = sum(y_2013), "2014" = sum(y_2014), "2015" = sum(y_2015), "2016" = sum(y_2016), "2017" = sum(y_2017), "2018" = sum(y_2018), "2019" = sum(y_2019), "2020" = sum(y_2020), "2021" = sum(y_2021)) %>%
  pivot_longer(cols = !group_name, names_to = "year", values_to = "value")

group_offenses_of_interest_sum
```
```{r}
group_offenses_of_interest_sum_2021 <- group_offenses_of_interest_sum %>%
  filter(year == "2021")

group_offenses_of_interest_sum_2021
```


```{r}
write_csv(group_offenses_of_interest_sum, "Data/Exports/april18_sum_group_offenses_of_interest.csv")
write_csv(group_offenses_of_interest_sum_2021, "Data/Exports/april21_group_offenses_of_interest_sum_2021.csv")
```



```{r}
year_count_table_view <- full_offenses_pd_approved %>%
  group_by(value_clean, clean_year) %>%
  tally() %>%
  arrange(clean_year) %>%
  pivot_wider(names_from = "clean_year", values_from = "n") %>%
  mutate(pchange_21_20 = ((y_2021 - y_2020)/y_2020)*100) %>%
  mutate(pchange_20_19 = ((y_2020 - y_2019)/y_2019)*100) %>%
  mutate(pchange_19_18 = ((y_2019 - y_2018)/y_2018)*100) %>%
  mutate(pchange_18_17 = ((y_2018 - y_2017)/y_2017)*100) %>%
  mutate(pchange_17_16 = ((y_2017 - y_2016)/y_2016)*100) %>%
  #ungroup() %>%
  arrange(desc(pchange_21_20)) %>%
  select(y_2011, y_2012, y_2013, y_2014, y_2015, y_2016, y_2017, y_2018, y_2019, y_2020, y_2021, pchange_21_20) %>%
  mutate_all(~replace(., is.na(.), 0)) %>%
  mutate(value_clean = str_to_title(value_clean))

year_count_table_view
```


## Analysis by Offenses

So there can be multiple offenses in an incident, looking at the offenses all together, here's what we know

First, we group the list by the sorted codes


```{r}
priority_year_grouping <- full_offenses_pd_approved %>%
  group_by(priority, year) %>%
  tally() %>%
  pivot_wider(names_from = priority, values_from = n) %>%
  clean_names() %>%
  mutate(total = (high+low+very_high+med)) %>%
  mutate(per_high = ((high/total)*100)) %>%
  mutate(per_low = ((low/total)*100)) %>%
  mutate(per_med = ((med/total)*100)) %>%
  mutate(per_v_high = ((very_high/total)*100)) %>%
  filter(year > as.Date("2015-01-01"))

priority_year_grouping
```
Cool, now we look at property crime

```{r}
property_crime_year_grouping <- full_offenses_pd_approved %>%
  group_by(property_crime, year) %>%
  tally() %>%
  pivot_wider(names_from = property_crime, values_from = n) %>%
  clean_names() %>%
  mutate(total = (y+n)) %>%
  mutate(per_yes_property_crime = ((y/total)*100)) %>%
  mutate(per_not_property_crime = ((n/total)*100)) 
  #filter(year > as.Date("2015-01-01"))

property_crime_year_grouping
```
##analysis by incident

this is a little bit harder. We need to look at each incident, which could have multiple different priorities associated with it and some non-property crimes associated with property crimes

So, I basically want to only look at incidents in this situation, so I will ungroup and narrow down my columns to remove any potential duplicates


** DELETE **

chief_approved_incident <- full_offenses_pd_approved %>%
  ungroup() %>%
  select(-desc_clean, -property_crime, -officer, -incident_month, -address_clean, -value_clean) %>%
  distinct()

chief_approved_incident

I count the number of incident numbers and create a situation in which the priority is seen as the tru priority if there is only one offense listed with the incident


incident_num <- chief_approved_incident %>%
  group_by(incident_num) %>%
  tally() %>%
  rename(count_incident_num = n)

chief_approved_incident <- chief_approved_incident %>%
  inner_join(incident_num, by = "incident_num") %>%
  mutate(tru_priority = case_when(
    count_incident_num == 1 ~ priority
  )) %>%
  select(-priority) %>%
  distinct()

chief_approved_incident


check_inc_num <- chief_approved_incident %>%
  group_by(incident_num) %>%
  tally() %>%
  arrange(desc(n))

check_inc_num


one_spec_inc <- chief_approved_incident %>%
  filter(incident_num == "11-00011(0)")

one_spec_inc

** DELETE **

i get ready to do some of my own prioritizing by grouping the incidents and their priorities and then counting them

```{r}
prep_by_incident <- full_offenses_pd_approved %>%
  #filter(count_incident_num >1) %>%
  group_by(incident_num, priority) %>%
  tally() %>%
  pivot_wider(values_from = n, names_from = priority) %>%
  mutate_all(~replace(., is.na(.), 0))
  

prep_by_incident
```

Now I create a system in which if there is a priority higher than other priorities listed, we result to the higher priority

```{r}
incident_real_priority <- prep_by_incident %>%
  clean_names() %>%
  mutate(tru_priority = case_when(
    low >0 & med >0 & high>0 & very_high >0 ~ "VERY HIGH",
    
    low >0 & high>0 & very_high >0 ~ "VERY HIGH",
    
    med >0 & high>0 & very_high >0 ~ "VERY HIGH",
    
    high > 0 & very_high >0 ~ "VERY HIGH",
    
    med > 0 & very_high >0 ~ "VERY HIGH",
    
    low > 0 & very_high >0 ~ "VERY HIGH",
    
    very_high != 0 & med == 0 & high == 0 & low == 0 ~ "VERY HIGH",
    
    med > 0 & high >0 ~ "HIGH", 
    
    low >0 & med >0 & high>0 ~ "HIGH",
    
    low > 0 & high >0 ~ "HIGH", 
    
    high != 0 & med == 0 & low == 0 & very_high == 0 ~ "HIGH",
    
    low > 0 & med >0 ~ "MED",
    
    med != 0 & low == 0 & high == 0 & very_high == 0 ~ "MED",
    
    low != 0 & med == 0 & high == 0 & very_high == 0 ~ "LOW",
    
    
    
  ))

incident_real_priority
```

trying to check myself here, is there any case without a priority assigned

```{r}
check_incident_priority <- incident_real_priority %>%
  filter(low >0 & med>0) 

check_incident_priority
```
Okay, glad I checked that

Moving forward to joining the incident to the tru priority

```{r}
final_chief_approved_by_incident <- full_offenses_pd_approved %>%
  ungroup() %>%
  select(-desc_clean, -property_crime, -officer, -incident_month, -address_clean, -value_clean, -priority) %>%
  distinct() %>%
  inner_join(incident_real_priority, by = "incident_num") %>%
  select(-low, -high, -med, -very_high)

final_chief_approved_by_incident
```
And here's a quick tally

```{r}
priority_year_grouping_incident <- final_chief_approved_by_incident %>%
  group_by(tru_priority, year) %>%
  tally() %>%
  pivot_wider(names_from = tru_priority, values_from = n) %>%
  clean_names() %>%
  mutate(total = (high+low+very_high+med)) %>%
  mutate(per_high = ((high/total)*100)) %>%
  mutate(per_low = ((low/total)*100)) %>%
  mutate(per_med = ((med/total)*100)) %>%
  mutate(per_v_high = ((very_high/total)*100)) 
  #filter(year > as.Date("2015-01-01"))

priority_year_grouping_incident
```



###incident: property crimes

** DELETE **

chief_approved_incident_property <- full_offenses_pd_approved %>%
  ungroup() %>%
  select(-desc_clean, -officer, -incident_month, -address_clean, -value_clean) %>%
  distinct()

chief_approved_incident_property



incident_num_prop <- chief_approved_incident_property %>%
  group_by(incident_num) %>%
  tally() %>%
  rename(count_incident_num = n)

chief_approved_incident_property <- chief_approved_incident_property %>%
  inner_join(incident_num, by = "incident_num") %>%
  select(-count_incident_num.x, -count_incident_num.y)

chief_approved_incident_property


** DELETE **



```{r}
prep_by_incident_prop <- full_offenses_pd_approved %>%
  #filter(count_incident_num >1) %>%
  group_by(incident_num, property_crime) %>%
  tally() %>%
  pivot_wider(values_from = n, names_from = property_crime) %>%
  ungroup() %>%
  mutate_all(~replace(., is.na(.), 0))

prep_by_incident_prop
```
  





```{r}
incident_real_property <- prep_by_incident_prop %>%
  #clean_names() %>%
  mutate(tru_property = case_when(
    Y > 0 & N == 0 ~ "Y",
    Y >0 & N>0 ~ "Y",
    N > 0 & Y == 0 ~ "N",
    
  ))


incident_real_property
```

```{r}
check_incident_real_property <- incident_real_property %>%
  filter(N >1 & Y>1) %>%
  group_by(tru_property) %>%
  tally()

check_incident_real_property
```



```{r}
final_chief_approved_by_incident_prop <- full_offenses_pd_approved %>%
  ungroup() %>%
  select(-desc_clean, -property_crime, -officer, -incident_month, -address_clean, -value_clean, -priority) %>%
  distinct() %>%
  inner_join(incident_real_property, by = "incident_num") %>%
  distinct()

final_chief_approved_by_incident_prop
```



```{r}
incident_property_crime_grouping <- final_chief_approved_by_incident_prop %>%
  group_by(tru_property, year) %>%
  tally() %>%
  pivot_wider(names_from = tru_property, values_from = n) %>%
  clean_names() %>%
  mutate(total = (y+n)) %>%
  mutate(per_yes_property_crime = ((y/total)*100)) %>%
  mutate(per_not_property_crime = ((n/total)*100)) 
  #filter(year > as.Date("2015-01-01"))

incident_property_crime_grouping
```


```{r}
write_csv(year_count, "Data/Exports/year_count.csv")
write_csv(year_count_table_view, "Data/Exports/year_count_table_view.csv")
#write_csv(property_crime_year_grouping, "Data/Exports/property_crime_year_grouping.csv")
#write_csv(priority_year_grouping, "Data/Exports/priority_year_grouping.csv")
write_csv(incident_property_crime_grouping, "Data/Exports/incident_property_crime_grouping.csv")
write_csv(priority_year_grouping_incident, "Data/Exports/priority_year_grouping_incident.csv")

#write_csv(codes_classified, "Data/Exports/Alamo_codes_2011_thru_2021_LS_labeled_categories.csv")

```

