---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(readxl)
```

Uploading the codes (That is the official police code list, approved by the chief.)

```{r}
codes_classified <- read_excel("Data/Exports/FOR REVIEW/Copy of march_14_codes_official.xlsx", sheet = 2)

codes_classified
```

```{r}
clean_codes_classified <- codes_classified %>%
  select(value_clean, priority, property_crime) %>%
  distinct() %>%
  filter(priority != "NA")

clean_codes_classified
```
```{r}
values_doubled_check <- clean_codes_classified %>%
  group_by(value_clean) %>%
  tally() %>%
  arrange(desc(n))

values_doubled_check
```

Perfect. Everything should be great

```{r}
full_offenses_pd_approved <- clean_offenses %>%
  inner_join(clean_codes_classified, by = "value_clean") %>%
  mutate(year = floor_date(date, unit = "year")) %>%
  mutate(clean_year = paste("y", substr(year, 1, 4), sep = "_")) %>%
  select(-rowid)

full_offenses_pd_approved
```
So, I'm going to use it to look at:

percent of each category of priority of crimes throughout the years
percent of property crimes throughout the years

First, this is just a general analysis of the crimes, what went up, what went down

```{r}
year_count <- full_offenses_pd_approved %>%
  group_by(value_clean, clean_year) %>%
  tally() %>%
  arrange(clean_year) %>%
  pivot_wider(names_from = "clean_year", values_from = "n") %>%
  mutate(pchange_21_20 = ((y_2021 - y_2020)/y_2020)*100) %>%
  mutate(pchange_20_19 = ((y_2020 - y_2019)/y_2019)*100) %>%
  mutate(pchange_19_18 = ((y_2019 - y_2018)/y_2018)*100) %>%
  mutate(pchange_18_17 = ((y_2018 - y_2017)/y_2017)*100) %>%
  mutate(pchange_17_16 = ((y_2017 - y_2016)/y_2016)*100) %>%
  arrange(desc(pchange_21_20)) %>%
  mutate(dif_21_20 = abs(y_2021-y_2020)) %>%
  filter(dif_21_20 > 15) %>%
  #filter(y_2020 >= 20) %>%
  select(value_clean, pchange_21_20, y_2020, y_2021, pchange_20_19, pchange_19_18, pchange_18_17, pchange_17_16, y_2019, y_2018, y_2017, y_2016, y_2015, y_2014, y_2013, y_2012, y_2011)

year_count
```
```{r}
year_count_table_view <- full_offenses_pd_approved %>%
  group_by(value_clean, clean_year) %>%
  tally() %>%
  arrange(clean_year) %>%
  pivot_wider(names_from = "clean_year", values_from = "n") %>%
  mutate(pchange_21_20 = ((y_2021 - y_2020)/y_2020)*100) %>%
  mutate(pchange_20_19 = ((y_2020 - y_2019)/y_2019)*100) %>%
  mutate(pchange_19_18 = ((y_2019 - y_2018)/y_2018)*100) %>%
  mutate(pchange_18_17 = ((y_2018 - y_2017)/y_2017)*100) %>%
  mutate(pchange_17_16 = ((y_2017 - y_2016)/y_2016)*100) %>%
  #ungroup() %>%
  arrange(desc(pchange_21_20)) %>%
  select(y_2011, y_2012, y_2013, y_2014, y_2015, y_2016, y_2017, y_2018, y_2019, y_2020, y_2021, pchange_21_20) %>%
  mutate_all(~replace(., is.na(.), 0)) %>%
  mutate(value_clean = str_to_title(value_clean))

year_count_table_view
```


## Analysis by Offenses

So there can be multiple offenses in an incident, looking at the offenses all together, here's what we know

First, we group the list by the sorted codes


```{r}
priority_year_grouping <- full_offenses_pd_approved %>%
  group_by(priority, year) %>%
  tally() %>%
  pivot_wider(names_from = priority, values_from = n) %>%
  clean_names() %>%
  mutate(total = (high+low+very_high+med)) %>%
  mutate(per_high = ((high/total)*100)) %>%
  mutate(per_low = ((low/total)*100)) %>%
  mutate(per_med = ((med/total)*100)) %>%
  mutate(per_v_high = ((very_high/total)*100)) %>%
  filter(year > as.Date("2015-01-01"))

priority_year_grouping
```
Cool, now we look at property crime

```{r}
property_crime_year_grouping <- full_offenses_pd_approved %>%
  group_by(property_crime, year) %>%
  tally() %>%
  pivot_wider(names_from = property_crime, values_from = n) %>%
  clean_names() %>%
  mutate(total = (y+n)) %>%
  mutate(per_yes_property_crime = ((y/total)*100)) %>%
  mutate(per_not_property_crime = ((n/total)*100)) 
  #filter(year > as.Date("2015-01-01"))

property_crime_year_grouping
```
##analysis by incident

this is a little bit harder. We need to look at each incident, which could have multiple different priorities associated with it and some non-property crimes associated with property crimes

So, I basically want to only look at incidents in this situation, so I will ungroup and narrow down my columns to remove any potential duplicates


** DELETE **

chief_approved_incident <- full_offenses_pd_approved %>%
  ungroup() %>%
  select(-desc_clean, -property_crime, -officer, -incident_month, -address_clean, -value_clean) %>%
  distinct()

chief_approved_incident

I count the number of incident numbers and create a situation in which the priority is seen as the tru priority if there is only one offense listed with the incident


incident_num <- chief_approved_incident %>%
  group_by(incident_num) %>%
  tally() %>%
  rename(count_incident_num = n)

chief_approved_incident <- chief_approved_incident %>%
  inner_join(incident_num, by = "incident_num") %>%
  mutate(tru_priority = case_when(
    count_incident_num == 1 ~ priority
  )) %>%
  select(-priority) %>%
  distinct()

chief_approved_incident


check_inc_num <- chief_approved_incident %>%
  group_by(incident_num) %>%
  tally() %>%
  arrange(desc(n))

check_inc_num


one_spec_inc <- chief_approved_incident %>%
  filter(incident_num == "11-00011(0)")

one_spec_inc

** DELETE **

i get ready to do some of my own prioritizing by grouping the incidents and their priorities and then counting them

```{r}
prep_by_incident <- full_offenses_pd_approved %>%
  #filter(count_incident_num >1) %>%
  group_by(incident_num, priority) %>%
  tally() %>%
  pivot_wider(values_from = n, names_from = priority) %>%
  mutate_all(~replace(., is.na(.), 0))
  

prep_by_incident
```

Now I create a system in which if there is a priority higher than other priorities listed, we result to the higher priority

```{r}
incident_real_priority <- prep_by_incident %>%
  clean_names() %>%
  mutate(tru_priority = case_when(
    low >0 & med >0 & high>0 & very_high >0 ~ "VERY HIGH",
    
    low >0 & high>0 & very_high >0 ~ "VERY HIGH",
    
    med >0 & high>0 & very_high >0 ~ "VERY HIGH",
    
    high > 0 & very_high >0 ~ "VERY HIGH",
    
    med > 0 & very_high >0 ~ "VERY HIGH",
    
    low > 0 & very_high >0 ~ "VERY HIGH",
    
    very_high != 0 & med == 0 & high == 0 & low == 0 ~ "VERY HIGH",
    
    med > 0 & high >0 ~ "HIGH", 
    
    low >0 & med >0 & high>0 ~ "HIGH",
    
    low > 0 & high >0 ~ "HIGH", 
    
    high != 0 & med == 0 & low == 0 & very_high == 0 ~ "HIGH",
    
    low > 0 & med >0 ~ "MED",
    
    med != 0 & low == 0 & high == 0 & very_high == 0 ~ "MED",
    
    low != 0 & med == 0 & high == 0 & very_high == 0 ~ "LOW",
    
    
    
  ))

incident_real_priority
```

trying to check myself here, is there any case without a priority assigned

```{r}
check_incident_priority <- incident_real_priority %>%
  filter(low >0 & med>0) 

check_incident_priority
```
Okay, glad I checked that

Moving forward to joining the incident to the tru priority

```{r}
final_chief_approved_by_incident <- full_offenses_pd_approved %>%
  ungroup() %>%
  select(-desc_clean, -property_crime, -officer, -incident_month, -address_clean, -value_clean, -priority) %>%
  distinct() %>%
  inner_join(incident_real_priority, by = "incident_num") %>%
  select(-low, -high, -med, -very_high)

final_chief_approved_by_incident
```
And here's a quick tally

```{r}
priority_year_grouping_incident <- final_chief_approved_by_incident %>%
  group_by(tru_priority, year) %>%
  tally() %>%
  pivot_wider(names_from = tru_priority, values_from = n) %>%
  clean_names() %>%
  mutate(total = (high+low+very_high+med)) %>%
  mutate(per_high = ((high/total)*100)) %>%
  mutate(per_low = ((low/total)*100)) %>%
  mutate(per_med = ((med/total)*100)) %>%
  mutate(per_v_high = ((very_high/total)*100)) 
  #filter(year > as.Date("2015-01-01"))

priority_year_grouping_incident
```



###incident: property crimes

** DELETE **

chief_approved_incident_property <- full_offenses_pd_approved %>%
  ungroup() %>%
  select(-desc_clean, -officer, -incident_month, -address_clean, -value_clean) %>%
  distinct()

chief_approved_incident_property



incident_num_prop <- chief_approved_incident_property %>%
  group_by(incident_num) %>%
  tally() %>%
  rename(count_incident_num = n)

chief_approved_incident_property <- chief_approved_incident_property %>%
  inner_join(incident_num, by = "incident_num") %>%
  select(-count_incident_num.x, -count_incident_num.y)

chief_approved_incident_property


** DELETE **



```{r}
prep_by_incident_prop <- full_offenses_pd_approved %>%
  #filter(count_incident_num >1) %>%
  group_by(incident_num, property_crime) %>%
  tally() %>%
  pivot_wider(values_from = n, names_from = property_crime) %>%
  ungroup() %>%
  mutate_all(~replace(., is.na(.), 0))

prep_by_incident_prop
```
  





```{r}
incident_real_property <- prep_by_incident_prop %>%
  #clean_names() %>%
  mutate(tru_property = case_when(
    Y > 0 & N == 0 ~ "Y",
    Y >0 & N>0 ~ "Y",
    N > 0 & Y == 0 ~ "N",
    
  ))


incident_real_property
```

```{r}
check_incident_real_property <- incident_real_property %>%
  filter(N >1 & Y>1) %>%
  group_by(tru_property) %>%
  tally()

check_incident_real_property
```



```{r}
final_chief_approved_by_incident_prop <- full_offenses_pd_approved %>%
  ungroup() %>%
  select(-desc_clean, -property_crime, -officer, -incident_month, -address_clean, -value_clean, -priority) %>%
  distinct() %>%
  inner_join(incident_real_property, by = "incident_num") %>%
  distinct()

final_chief_approved_by_incident_prop
```



```{r}
incident_property_crime_grouping <- final_chief_approved_by_incident_prop %>%
  group_by(tru_property, year) %>%
  tally() %>%
  pivot_wider(names_from = tru_property, values_from = n) %>%
  clean_names() %>%
  mutate(total = (y+n)) %>%
  mutate(per_yes_property_crime = ((y/total)*100)) %>%
  mutate(per_not_property_crime = ((n/total)*100)) 
  #filter(year > as.Date("2015-01-01"))

incident_property_crime_grouping
```


```{r}
write_csv(year_count, "Data/Exports/year_count.csv")
write_csv(year_count_table_view, "Data/Exports/year_count_table_view.csv")
#write_csv(property_crime_year_grouping, "Data/Exports/property_crime_year_grouping.csv")
#write_csv(priority_year_grouping, "Data/Exports/priority_year_grouping.csv")
write_csv(incident_property_crime_grouping, "Data/Exports/incident_property_crime_grouping.csv")
write_csv(priority_year_grouping_incident, "Data/Exports/priority_year_grouping_incident.csv")

#write_csv(codes_classified, "Data/Exports/Alamo_codes_2011_thru_2021_LS_labeled_categories.csv")

```

